---
title: "Covid91 vaccine study Final2024"
author: "Mya Orozco"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Fall 2024`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
Males <- guy %>%
  group_by(infected,treatment) %>%
  summarize(count=n()) %>% 
  mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This chart gives the number of guys(males) and how they are related to COVID-19 by either taking the vaccine or placebo.


The plot shows that covid19 is much lower in people who took the vaccine than in those who got the placebo. The size of the placebo group compared to the vaccine group is about 3:1.


##Numerical results


```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

It is shown that even though the vaccine was given to males, a relatively lower percentage, 26.2%, contracted COVID-19. On the contrary, data shows that many more, 73.8%, of those who did not take any of the vaccines eventually contracted COVID-19. This comparison truly shows that the vaccine proves to be very promising and effective against the virus.



### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```


The information that is really important to pay attention to out of all these findings is the output of the Fisher's Exact test, which gives us some significant findings. The p-value, which is shown above, is very small and stands at 2.2e-16, which can also be expressed as 0.00000000000000022. A number this small strongly suggests that such results are not easily obtainable by chance alone; hence, there must be an association.

Our odds ratio of approximately 2.8 proves that this vaccine is effective in stopping covid19 compared to a placebo.



```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`

The odds ratio for males getting vaccines compared to placebo is well over 70. That would indicate we have a very effective vaccine for use in males.



## Females

### Barplot

```{r}
Females <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of gals(females) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo.

The Covid19 vaccine has been shown to have a 2:1 ratio regarding the number of people infected when this is compared to those who have been given the placebo treatment. In particular, the percent of females who became infected following the receipt of the vaccine seems to be about 36%, while, in contrast, the percent of females who were infected following the receipt of the placebo is about 65%


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

The females who got the vaccine, 35% went on to get COVID-19. In comparison, of those who were given the placebo, a much larger 65% became infected with the virus. These stats show a pretty large disparity, illustrating a ratio roughly 2:1 for the vaccinated versus placebo groups.


































## LGBTQ



## Druggies


# Overall Results and Conclusions